MA_Results

### 2024-03-19 Markdown/Manuscript/Child_RMDs/MS_Imp_Guidelines.Rmd
v4

<!-- We considered literature on modelling practice and reflected on our prior experience with developing and using CHEMs to identify: (i) some core ethical responsibilities of CHEM project teams; (ii) attributes of CHEMs that can suggest fulfillment of these responsibilities; and (iii) criteria against which these attributes can be assessed.  -->
<!-- responsibilities to goals -->
# Ethical goals for CHEM developers and funders
Boden and McKendrick [@10.3389/fpubh.2017.00068] propose a framework for ethical public health modelling based on the criteria of independence (concerning how modeller subjectivity shapes model design), beneficence / non-maleficence (concerning model quality and utility), transparency (concerning the need for policymakers to reliably evaluate model strengths and weaknesses) and justice (concerning the social obligations of modellers to consider and communicate ethical issues about model use). <!-- However, t -->These criteria are not specific to the computational model, but <!-- are designed to  -->apply to a modelling project as a whole, including a model's conceptual and mathematical representations. The authors propose 13 questions to evaluate ethical risk across the four criteria, but of these, only "is the model code open source or available on request" (for transparency) and "has the model been verified, i.e., does it do what the modeller wants it to do?" (for beneficence / non-maleficence) are specific to model's representation as software.

To identify ethical issues that are specific to CHEMs, we <!-- first  -->considered <!-- Edit -->how Boden and McKendrick's four criteria might apply to <!-- /Edit -->the distinct phases in a software development project. Software development lifecycle (SDLC) models take different approaches to the inclusion, naming, definition, sequencing and iteration of project phases [@ruparelia2010software]. However, SDLC models typically have components that can principally map to the concepts of *planning* (e.g., identification of system / user requirements and design specifications), *implementation* (e.g., development and testing) and *release* (e.g., deployment, system integration, maintenance and support). For simplicity, we propose one overarching ethical <!-- responsibility  for CHEM project team-->goal for each phase. These goals<!-- responsibilities  --> are: 
<!-- establish  -->
(i) *socially acceptable user requirements and design specifications* (during CHEM planning);
<!-- ensure  -->
(ii) *fitness for purpose* (of CHEM implementations); and
<!-- support  -->
(iii) *socially beneficial use* (post CHEM release).

Boden and McKendrick’s transparency criterion underpins the ethical goals we propose for all three phases (Table \@ref(tab:timelygls)). Boden and McKendrick’s other <!-- ethical  -->criteria can be selectively mapped to specific phases: our proposed goal for ethical CHEM planning incorporates justice and independence considerations; the goal we propose for CHEM implementation is based on independence and beneficence / non-maleficence criteria; and our proposed ethical goal post CHEM release reflects beneficence / non-maleficence and justice criteria. <!-- /Edit -->

Misalignment between the values of computational model developers and those of the population groups affected by decisions based on their models presents significant ethical risks [@thompson2022escape; @duckett2022journey].  The value judgments of health economic modellers are rarely adequately specified [@duckett2022journey]. These value judgments influence the assumptions, selection of model features and standards for evidence that shape health economic model projects [@HARVARD2020112975]. For example, the process of determining CHEM *user requirements and design specifications* will involve making judgments about the competing claims of alternative software development platforms (Table \@ref(tab:proscons)). That decision will in turn determine how straightforward it is for a CHEM to be subsequently amended to reflect alternative value judgments relating to model conceptualisation and mathematical formalisation (Table \@ref(tab:timelygls)).
<!-- Moved -->
<!--  , omissions that may lead to socially unacceptable policy recommendations -->

Modellers need to do more than self-certify that their <!-- have duties both to ensure a  -->computational model is *fit for purpose* - <!-- Edit -->they also need to<!-- /Edit --> provide potential third-party users with the means of assessing its adequacy 
<!-- for their proposed uses  -->[@Erdemir2020; @Feenstra2022; @thompson2019escape; @10.3389/fpubh.2017.00068]. However, it is common for health economic models to have serious methodological flaws [@carletto_zanuzzi_sammarco_russo_2020; @WONDER2015467]; insufficient validation [@Ghabri2019; @kolovos2017model; @haji2013model], poor reproducibility [@Jalali2021; @McManus2019; @Bermejo2017]; and undeclared errors [@Radeva2020]. Appropriate computational implementation choices can help address many of these shortcomings, for example by automating quality assurance checks and <!-- Edit -->providing executable code for  <!--facilitating manual  -->review and testing <!-- /Edit -->by third parties.

The *social benefit* of developing a CHEM may be limited if it is rarely used, if mis-used or when its acceptability and adequacy rapidly decay. Reuse of CHEMs as components of other models can potentially make model development more efficient [@Arnold2010; @garcia2021cost]. However, health economic models face challenges related to transferability across jurisdictions [@garcia2021cost] that create barriers to reuse. Without ongoing maintenance, a CHEM risks becoming less reliable with time [@garcia2021cost] and is at risk of being deployed for purposes for which it is poorly suited [@calder2018computational]. Currently, health economic models are rarely implemented computationally in a manner that facilitates routine updates [@Sampson2017], thus limiting the temporal window within which a CHEM can be validly applied. 
<!-- Moved -->
<!-- Edit --><!-- All three  -->Our proposed ethical goals are most likely to be met when there is both good<!-- have implications for both  --> professional practice by CHEM developers and <!-- Addition -->adequate <!-- Addition -->resourcing by project funders.  <!-- /Edit --> The choices CHEM developers make often need to balance ethical considerations with project resource constraints (as illustrated in Table \@ref(tab:proscons)). Modelling projects implemented by diligent and conscientious CHEM developers may still be exposed to ethical risks if there are inadequate resources to meaningfully engage important stakeholder groups, prepare code and data artefacts for independent assessment and provide ongoing model maintenance and user support. Existing incentive structures for research health economists generally do not promote facilitating peers to reuse their work and funding is rarely adequate to maintain and routinely update a model. Currently, it can take “an extraordinary amount of idealism” to commit to authoring and maintaining research software [@anzt2020environment]. For these reasons, we believe responsibility for achieving our proposed ethical goals is shared between CHEM developers and funders. Reflecting the shared nature of this responsibility, we suggest that in most cases where a modelling project fails to meet our ethical goals, it will be more appropriate to describe these shortcomings as "ethical risks" (the language used by Boden and McKendrick) than to use accusatory terms like "unethical practice."<!-- these responsibilities should apply jointly to health economic model developers and funders. -->

@@

Taking steps to make computational models *transparent* is an important means of reducing ethical risk <!-- part of ethical public health modelling practice  -->[@10.3389/fpubh.2017.00068]. Guidance on transparency in health economic modelling recommend that model code and data should be clearly documented [@Eddy2012]. Notably, the same guidelines, published over ten years ago, do not recommend sharing model code and data. However, more recent healthcare modelling guidance does recommend public dissemination of such artefacts [@Erdemir2020]. Online repository services such as Zenodo [@Zenodo2013] and Dataverse [@Dataverse2007] provide persistent storage solutions that generate a Digital Object Identifier (DOI) for each code and data collection. Ensuring that calculations are correct and consistent with model specifications is an essential part of CHEM quality assurance [@techver2019]. The extensiveness of such verification checks can be reported using the concept of code coverage [@ERICWONG2010188] - the proportion of model code that has been explicitly tested. Tests should ideally combine both unit tests (to verify that small, isolated sections of code produce the correct output when run independently) and acceptance tests (to verify that the correct output is produced when multiple code components are run together to perform tasks that meet core user-requirements [@martin2003agile]). The nature and extent of individual model authorship contributions can become unclear when models are implemented over longer time-frames with a large and changing group of collaborators [@thompson2022escape]. This issue can be addressed by use of online repository services such as GitHub [@github2007], that provide citation tools and can transparently record all individual code contributions to a modelling project over its lifecycle.

<!-- A CHEM that is *reusable* also signals ethical modelling practice.  -->
Although making a CHEM’s code, data and documentation publicly available is increasingly considered good practice, it is insufficient to ensure that a model is *reusable*. Key concerns for health economists when considering whether to reuse a model are generalisability (application without adaptation) and transferability (selective reuse and/or modification of model components) [@RN39]. Writing model algorithms as collections of functions (short, self-contained and reusable software routines that each perform a discrete task) is good scientific computing practice [@Wilson_2017] and promotes selective reuse. Computational implementations that store model code and data in distinct files and locations (as opposed to embedding data such as parameter values into source code) are also easier to selectively modify. Modular implementations enable models to be constructed from multiple independently reusable and replaceable sub-models (modules) [@pan2021modular]. The programming concept inheritance [@8717448], deployed in object-oriented programming approaches, can facilitate the selective reuse and modification of model components. Granting permissions to others to use, test and adapt models and their components, can be facilitated by two broad categories of open source licenses. Some guidance strongly recommends the use of permissive licensing [@Wilson_2017] that provides users with great flexibility as to the purposes (including commercial) for which content can be re-used. Alternatively, copyleft licenses [@copyleft2022] can be used to require that derivative works created by content users remain open-source.

Models should be *updatable* so that they remain valid for longer, evolving as new evidence emerges and the systems being modelled change [@Jenkins2021; @garcia2021cost]. Ensuring that a model is regularly reviewed to identify and implement required improvements is a recommended defence against model validity decay [@calder2018computational]. Sustainable maintenance of open source research software requires both a core development team and an active user community [@info:doi/10.2196/20028]. Online communities can be an efficient means of engaging model users in testing each version of a model, identifying issues and suggesting improvements. Services such as GitHub [@github2007] provide collaborative code development tools [@MERGEL2015464] that help elicit, integrate and reconcile contributions from multiple contributors and to ensure each update is uniquely identifiable and retrievable. It is also important that verification checks are rerun with each model update, a task that can be automated using the software development practice of continuous integration [@CI2017].  The risk of software revisions having unintended consequences for third party users can be mitigated through the use of deprecation conventions [@zhou2016api] that take an informative and staged approach to retiring outdated CHEMs.

<!-- # -->
#	Criteria for assessing TRU attributes
We suggest six criteria (two for each attribute) against which for assessing the TRU attributes of CHEMs.

How *transparent* a CHEM is can be assessed against the criteria:

-	*T1*: Software files (executable code, testing procedures and outcomes and non-confidential data) are available in open access repositories.

-	*T2*: It is easy to see who funded, developed and tested each part of the CHEM and to identify the modelling team’s assumptions, judgments and theories about CHEM development and use.

How *reusable* a CHEM is can be assessed against the criteria:

-	*R1*: Programming practices facilitate selective reuse of model sub-components.<!-- Replacement -->

-	*R2*: Licenses restrict only unethical use and allow anyone to ethically reuse model code and non-confidential data, in whole or in part, without charge, and for purposes that include the creation of derivative works.

How *updatable* a CHEM is can be assessed against the criteria:

-	*U1*: Maintenance infrastructure is in place to support version control and collaboration with model users.

-	*U2*: Each new release of a CHEM is systematically retested, with changes implemented to minimise disruptions for existing model users.

>>>>

<!-- We considered literature on modelling practice and reflected on our prior experience with developing and using CHEMs to identify: (i) some core ethical responsibilities of CHEM project teams; (ii) attributes of CHEMs that can suggest fulfillment of these responsibilities; and (iii) criteria against which these attributes can be assessed.  -->
<!-- responsibilities to goals -->
# Research ethics goals for CHEM developers and funders
Resnik [@resnik2020ethics] identified eighteen ethical principles that underpin influential codes of research good practice. We suggest that six of these principles are particularly relevant to CHEM implementation decisions: *accountability* (researchers declaring and taking responsibility for their specific contributions); *confidentiality* (protecting patient records and other confidential information), *intellectual property* (honoring copyrights and acknowledging contributors), *openness* (sharing project data, tools and resources), *transparency* (disclosing the methods and materials needed to evaluate research) and *social responsibility* (promoting social good).

Boden and McKendrick [@10.3389/fpubh.2017.00068] propose a framework for ethical public health modelling based on the criteria of *independence* (concerning how modeller subjectivity shapes model design), *beneficence / non-maleficence* (concerning model quality and utility), *transparency* (concerning the need for policymakers to reliably evaluate model strengths and weaknesses) and *justice* (concerning the social obligations of modellers to consider and communicate ethical issues about model use). <!-- However, t -->This framework is not specific to a model's implementation as software, but <!-- are designed to  -->apply to a modelling project as a whole, including a model's conceptual and mathematical representations. Boden and McKendrick proposed 13 questions to evaluate ethical risk across the four criteria, but of these, only two pertain to computational models: "is the model code open-source or available on request" (for transparency) and "has the model been verified, i.e., does it do what the modeller wants it to do?" (for beneficence / non-maleficence).

To identify ethical issues that are specific to CHEMs, we <!-- first  -->considered <!-- Edit -->how Resnick's 18 principles and Boden and McKendrick's four criteria might apply to <!-- /Edit -->the distinct phases in a software development project. Software development lifecycle (SDLC) models take different approaches to the inclusion, naming, definition, sequencing and iteration of project phases [@ruparelia2010software]. However, SDLC models typically have components that can also principally map to the concepts of *planning* (e.g., identification of system / user requirements and design specifications), *implementation* (e.g., development and testing) and *release* (e.g., deployment, system integration, maintenance and support). For simplicity, we propose one overarching ethical <!-- responsibility  for CHEM project team-->goal for each phase. These goals<!-- responsibilities  --> are: 
<!-- establish  -->
(i) *socially acceptable user requirements and design specifications* (during CHEM planning); 
<!-- ensure  -->
(ii) *fitness for purpose* (of CHEM implementations); and 
<!-- support  -->
(iii) *socially beneficial use* (post CHEM release).

Transparency (as a Resnik principle and Boden and McKendrick criterion) and Resnik's social responsibility principle underpin all three of the ethical goals we propose (Table \@ref(tab:timelygls)). Additionally: <!-- Boden and McKendrick’s other <!-- ethical  --><!-- criteria can be selectively mapped to specific goals:  -->(i) the social acceptability goal incorporates Resnik's accountability, confidentiality and intellectual property principles and Boden and McKendrick's justice and independence criteria; 
(ii) the fitness for purpose goal relates to Resnik's accountability principle and Boden and McKendrick's independence and beneficence / non-maleficence criteria; and 
(iii) the socially beneficial use goal reflects Resnik's openness principle and Boden and McKendrick's beneficence / non-maleficence and justice criteria. <!-- /Edit -->

Misalignment between the values of computational model developers and those of the population groups affected by decisions based on their models presents significant ethical risks [@thompson2022escape; @duckett2022journey].  The value judgments of health economic modellers are rarely adequately specified [@duckett2022journey]. These value judgments influence the assumptions, selection of model features and standards for evidence that shape health economic model projects [@HARVARD2020112975]. For example, the process of determining CHEM *user requirements and design specifications* will involve making judgments about the competing claims of open-source and commercial development platforms (Table \@ref(tab:proscons)). That decision will in turn determine how straightforward it is for a CHEM to be subsequently amended to reflect alternative value judgments relating to model conceptualisation and mathematical formalisation (Table \@ref(tab:timelygls)).
<!-- Moved -->
<!--  , omissions that may lead to socially unacceptable policy recommendations -->

Modellers need to do more than self-certify that their <!-- have duties both to ensure a  -->computational model is *fit for purpose* - <!-- Edit -->they also need to<!-- /Edit --> provide potential third-party users with the means of assessing its adequacy, appropriateness and usefulness <!-- for their proposed uses  -->[@Erdemir2020; @Feenstra2022; @thompson2019escape; @10.3389/fpubh.2017.00068]. However, it is common for health economic models to have serious methodological flaws [@carletto_zanuzzi_sammarco_russo_2020; @WONDER2015467]; insufficient validation [@Ghabri2019; @kolovos2017model; @haji2013model], poor reproducibility [@Jalali2021; @McManus2019; @Bermejo2017]; and undeclared errors [@Radeva2020]. Appropriate computational implementation choices can help address many of these shortcomings, for example by automating and publicly documenting quality assurance checks and <!-- Edit -->providing executable code for  <!--facilitating manual  -->review and testing <!-- /Edit -->by third parties.

The *social benefit* of developing a CHEM may be limited if it is rarely used, if mis-used or when its acceptability and adequacy rapidly decay. Reuse of CHEMs as components of other models can potentially make model development more efficient [@Arnold2010; @garcia2021cost]. However, health economic models face challenges related to transferability across jurisdictions [@garcia2021cost] that create barriers to reuse. Without ongoing maintenance, a CHEM risks becoming less reliable with time [@garcia2021cost] and is at risk of being deployed for purposes for which it is poorly suited [@calder2018computational]. Currently, health economic models are rarely implemented computationally in a manner that facilitates routine updates [@Sampson2017], thus limiting the temporal window within which a CHEM can be validly applied.

<!-- Moved -->
<!-- Edit --><!-- All three  -->Our proposed ethical goals are most likely to be met when there is both good<!-- have implications for both  --> professional practice by CHEM developers and <!-- Addition -->adequate <!-- Addition -->resourcing by project funders. <!-- /Edit --> The choices CHEM developers make often need to balance ethical considerations with project resource constraints (as illustrated in Table \@ref(tab:proscons)). Modelling projects implemented by diligent and conscientious CHEM developers may still be exposed to ethical risks if there are inadequate resources to meaningfully engage important stakeholder groups, prepare code and data artefacts for independent assessment and provide ongoing model maintenance and user support. Currently, it can take “an extraordinary amount of idealism” to commit to authoring and maintaining research software [@anzt2020environment]. Existing incentive structures for research health economists generally do not promote facilitating peers to reuse their work and funding is rarely adequate to maintain and routinely update a model. Economists may therefore make pragmatic judgments that proprietary licensing approaches provide the most feasible pathway to mobilising the resources to support model maintenance. For these reasons, we believe responsibility for achieving our proposed ethical goals is shared between CHEM developers and funders. Reflecting the shared nature of this responsibility, we suggest that in most cases where a modelling project fails to meet our ethical goals, it will be more appropriate to describe these shortcomings as "ethical risks" (the language used by Boden and McKendrick) than to use accusatory terms like "unethical practice."<!-- these responsibilities should apply jointly to health economic model developers and funders. -->

@@

Taking steps to make computational models *transparent* is an important means of reducing ethical risk <!-- part of ethical public health modelling practice  -->[@10.3389/fpubh.2017.00068]. Guidance on transparency in health economic modelling recommend that model code and data should be clearly documented [@Eddy2012]. These transparency guidelines, published twelve years ago, did not recommend sharing model code and data. However, more recent healthcare modelling guidance does recommend public dissemination of such artefacts [@Erdemir2020]. Online repository services such as Zenodo [@Zenodo2013] and Dataverse [@Dataverse2007] provide persistent storage solutions that generate a Digital Object Identifier (DOI) for each code and data collection. Ensuring that calculations are correct and consistent with model specifications is an essential part of CHEM quality assurance [@techver2019]. The extensiveness of such verification checks can be reported using the concept of code coverage [@ERICWONG2010188] - the proportion of model code that has been explicitly tested. <!-- Tests should ideally combine both unit tests (to verify that small, isolated sections of code produce the correct output when run independently) and acceptance tests (to verify that the correct output is produced when multiple code components are run together to perform tasks that meet core user-requirements [@martin2003agile]). -->Testing procedures and results can be publicly shared through online code development services like GitHub [@github2007]. GitHub also provides citation tools and can transparently record all individual code contributions to a modelling project over its lifecycle. These tools may be particularly useful in complex modelling projects implemented over longer time-frames with a large and changing group of collaborators, where the nature and extent of individual model authorship contributions can become unclear [@thompson2022escape].

<!-- A CHEM that is *reusable* also signals ethical modelling practice.  -->
Although making a CHEM’s code, data and documentation publicly available is increasingly considered good practice, it is insufficient to ensure that a model is *reusable*. <!-- Key concerns for health economists when considering whether to reuse a model are generalisability (application without adaptation) and transferability (selective reuse and/or modification of model components) [@RN39].  -->Writing model algorithms as collections of functions (short, self-contained and reusable software routines that each perform a discrete task) has been recommended as good scientific computing practice [@Wilson_2017] and promotes reuse of model sub-components. Computational implementations that store model code and data in distinct files and locations (as opposed to embedding data such as parameter values into source code) are also easier to selectively modify. Modular implementations enable models to be constructed from replaceable sub-models (modules) [@pan2021modular] that can share inputs and outputs with each other or be run as independent models [@barros2023empowering]. Modularity can <!-- Advantages of modular models include feasibility (large projects are broken into smaller tasks, with each component independently developed and tested) and flexibility ( -->make it easier to selectively replace or update specific parts of a model and to scale up or down the level of granularity [@pan2021modular]. <!-- The programming concept inheritance [@8717448], deployed in object-oriented programming approaches, can facilitate the selective reuse and modification of model components.  -->Granting permissions to others to use, test and adapt models and their components, can be facilitated by two broad categories of open-source licenses. Some guidance strongly recommends the use of permissive licensing [@Wilson_2017] that provides users with great flexibility as to the purposes (including commercial) for which content can be re-used. Alternatively, copyleft licenses [@copyleft2022] can be used to require that derivative works created by content users remain open-source.

Models should be *updatable* so that they remain valid for longer, evolving as new evidence emerges and the systems being modelled change [@Jenkins2021; @garcia2021cost]. Ensuring that a model is regularly reviewed to identify and implement required improvements is a recommended defence against model validity decay [@calder2018computational]. Sustainable maintenance of open-source research software requires both a core development team and an active user community [@info:doi/10.2196/20028]. Online communities can be an efficient means of engaging model users in testing each version of a model, identifying issues and suggesting improvements. Services such as GitHub [@github2007] provide collaborative code development tools [@MERGEL2015464] that help elicit, integrate and reconcile contributions from multiple contributors and to ensure each update is uniquely identifiable and retrievable. Verification checks should be rerun with each model update, a task that can be automated using the software development practice of continuous integration [@CI2017].  The risk of software revisions having unintended consequences for third party users can be mitigated through the use of deprecation conventions [@zhou2016api] that take an informative and staged approach to retiring outdated CHEMs.

<!-- # -->
#	Criteria for assessing TRU attributes
We suggest six criteria (two for each attribute) for assessing the TRU attributes of CHEMs. These criteria are summarised in Table \@ref(tab:trucrit).

```{r, child = child_docs_ls$three, echo=FALSE}
```

<!-- How *transparent* a CHEM is can be assessed against the criteria: -->

<!-- -	*T1*: Software files (executable code, testing procedures and outcomes and non-confidential data) are publicly available. -->

<!-- -	*T2*: It is easy to see who funded, developed and tested each part of the CHEM and to identify the modelling team’s assumptions, judgments and theories about CHEM development and use. -->

<!-- How *reusable* a CHEM is can be assessed against the criteria: -->

<!-- -	*R1*: Programming practices facilitate selective reuse of model sub-components.<!-- Replacement -->

<!-- -	*R2*: Licenses allow anyone to ethically reuse model code and non-confidential data, in whole or in part, without charge, and for purposes that include the creation of derivative works. -->

<!-- How *updatable* a CHEM is can be assessed against the criteria: -->

<!-- -	*U1*: Maintenance infrastructure is in place to support version control and collaboration with model users. -->

<!-- -	*U2*: Each new release of a CHEM is systematically retested, with changes implemented to minimise disruptions for existing model users. -->

### 2024-03-19 Markdown/Manuscript/Child_RMDs/MS_Intro_Short.Rmd
for coauthor feedback draft
v4

Health economics is a discipline concerned with problems that arise due to scarce resources, such as how to value health and healthcare, allocate healthcare budgets and configure health services [@wagstaff2012four]. In seeking to solve these problems, health economists typically use models which are simplified and selective representations of systems that are believed to influence human health. A health economic model should be capable of being described using words and figures (a conceptual model), equations (a mathematical model) and software (a computational model).  Health economic scientific manuscripts typically describe a model using its conceptual and mathematical representations but report results that have been generated by its computational representation (i.e., the format that allows computers to perform calculations). The conceptual, mathematical and computational representations of a model are assumed to be internally consistent. However, the ease of independently assessing the validity of this assumption depends in part on how a health economic model has been implemented computationally.

The set of software files implementing a health economic model's structure, data and algorithms can be called a computational health economic model (CHEM). CHEMs can be developed using commercial modelling <!-- Addition -->or spreadsheet<!-- /Addition --> software or open-source programming languages such as Python [@python2009] or R [@RCORE2022]. An advantage of commercial <!-- modelling  -->software is that users often require limited or no programming skills to develop and apply CHEMs. <!-- Moved -->Open-source CHEMs remain rare [@Feenstra2022; @Emerson2019]<!-- Moved -->. However, <!-- Addition -->it has been argued that<!-- /Addition --> open-source software programming languages may facilitate the development of CHEMs that are more transparent, reusable and updatable [@incerti2019r; @Pouwels2022]. The decision about what type of software development platform to use must balance ethical considerations and resource constraints. <!-- ( see--><!-- Addition --> In Table \@ref(tab:proscons), we summarise some of the trade-offs we consider when undertaking modelling projects<!-- ) -->.

@@

<!-- A high-level software framework for implementing open-source CHEMs in R has previously been developed with the primary aim of improving model transparency [@alarid2019need]. However, to gain widespread adoption, a software framework for CHEMs may need to provide tools to help model developers meet a broader range of ethical practice requirements. There are multiple challenges to overcome. -->
<!-- Edit Move -->Determining how to make these trade offs can be difficult. The computational modelling landscape is changing fast [@calder2018computational] and the ethical <!-- responsibilities  --><!-- Edit -->dimensions of computational public health modelling <!-- /Edit -->remain poorly understood [@10.3389/fpubh.2017.00068]. <!-- /Edit Move-->

>>>>

Health economics is a discipline concerned with problems that arise due to scarce resources, such as how to value health and healthcare, allocate healthcare budgets and configure health services [@wagstaff2012four]. In seeking to solve these problems, health economists typically use models which are simplified and selective representations of systems that are believed to influence human health. A health economic model should be capable of being described using words and figures (a conceptual model), equations (a mathematical model) and software (a computational model).  Health economic scientific manuscripts typically involve description of a model using its conceptual and mathematical representations and report results that have been generated by its computational representation (i.e., the format that allows computers to perform calculations). Authors of these manuscripts assume (rarely explicitly) that the conceptual, mathematical and computational representations of a model are internally consistent and externally valid. The feasibility of independently assessing the validity of this assumption depends in part on how a health economic model has been implemented computationally.

The set of software files implementing a health economic model's structure, data and algorithms can be called a computational health economic model (CHEM). CHEMs can be developed using commercial modelling <!-- Addition -->or spreadsheet<!-- /Addition --> software (for example, Microsoft Excel [@msexcel]) or open-source programming languages such as Python [@python2009] or R [@RCORE2022]. An advantage of commercial <!-- modelling  -->software is that users often require limited or no programming skills to develop and apply CHEMs.  <!-- Addition -->It has been argued that<!-- /Addition --> open-source software programming languages may facilitate the development of CHEMs that are more transparent, reusable and updatable [@incerti2019r; @Pouwels2022]. <!-- Moved -->However, open-source CHEMs remain rare [@Feenstra2022; @Emerson2019]<!-- Moved -->.

Health economists must balance ethical considerations and resource constraints when deciding what type of software development platform to use for authoring CHEMs. <!-- ( see--><!-- Addition --> In Table \@ref(tab:proscons), we summarise some of the trade-offs we consider when undertaking modelling projects<!-- ) -->. <!-- Edit Move -->Determining how to make these trade-offs can be difficult. The computational modelling landscape is changing fast [@calder2018computational] and the ethical <!-- responsibilities  --><!-- Edit -->dimensions of computational public health modelling <!-- /Edit -->remain poorly understood [@10.3389/fpubh.2017.00068]. <!-- /Edit Move-->

@@

<!-- A high-level software framework for implementing open-source CHEMs in R has previously been developed with the primary aim of improving model transparency [@alarid2019need]. However, to gain widespread adoption, a software framework for CHEMs may need to provide tools to help model developers meet a broader range of ethical practice requirements. There are multiple challenges to overcome. -->

### 2024-03-19 Markdown/Manuscript/Child_RMDs/MS_Tables.Rmd
v4

```{r echo=FALSE}
if(params$X@tables_in_body_lgl){
  child_docs_ls <- list(one = "../Child_RMDs/MS_Tables_1.Rmd",
                        two = "../Child_RMDs/MS_Tables_2.Rmd"
                        # ,
                        # three = "../Child_RMDs/MS_Tables_3.Rmd",
                        # four = "../Child_RMDs/MS_Tables_4.Rmd"
                        )
}else{
    child_docs_ls <- list(one = NULL,
                          two = NULL,
                          three = NULL,
                          four = NULL)
}
```

>>>>

```{r echo=FALSE}
if(params$X@tables_in_body_lgl){
  child_docs_ls <- list(one = "../Child_RMDs/MS_Tables_1.Rmd",
                        two = "../Child_RMDs/MS_Tables_2.Rmd",
                        three = "../Child_RMDs/MS_Tables_3.Rmd"
                        #,
                        # four = "../Child_RMDs/MS_Tables_4.Rmd"
                        )
}else{
    child_docs_ls <- list(one = NULL,
                          two = NULL,
                          three = NULL,
                          four = NULL)
}
```
